### 2019-06-04 Documents/literature_review/pages/review_content.tex
Changes; - Added more content and worked on the wording of the paper

\subsection*{Encoding scheme performance}
Results 
from the literature consistently show the effectiveness of language based encodings (Words or Sentences). This was shown in \cite{dechand2016empirical}, \cite{tan2017can} and \cite{kainda2009usability} with the accuracy of these schemes all being above 94.00\%. In all cases these were the best scemes in terms of accuracy. The exception to this is the work performed by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study} in 2009 with Words achieving an abnormal accuracy of 63.00\% (Comparisons of methodology will be discussed in later sections).
\\\\
Aside from textual representations were graphical schemes. Examples of schemes assessed were: Random Art\cite{perrig1999hash}, Flag\cite{ellison2003public}, T-Flag\cite{lin2010spate}, Vash\footnote{https://github.com/thevash/vash}, OpenSSH visual host key and Unicorns\footnote{https://unicornify.pictures/} to name a few. Accuracy from these schemes were mixed with positive results obtained by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study} where their best accuracy was achived by Random Art (94.00\%) and Flag Ext.\footnote{The authors' own improvement on Flag}(88.00\%). On the other hand, the worst accuracy was obtained by Flag(50\%). The only other paper assessing graphical representations was the work of \textbf{Tan \textit{et al.}}\cite{tan2017can} where they also achieved mixed results with accuracies ranging from 46\% to 90\%.
\\
In terms of usability of graphical schemes the literature concurred on their ease-of-use. In terms of speed of comparison all schemes consistently obtained some of the quickest comparisons times. This is also supported indirectly from the the comparison results of alternative literature \cite{kainda2009usability}\cite{dechand2016empirical} where with these in consideration graphical schemes still remain the quickest overall.
In terms of research into the performance of graphical schemes, the literature does not contain an extensive review with only two papers including graphical schemes in their comparison. This is, therefore, a candidate for further research.

One unique paper in this research area was the work by \textbf{M. Shirvanian \textit{et al.}}\cite{shirvanian2017pitfalls} produced in the context of secure messaging pairing. This paper was unique for a number of reasons. First was to consideration for ``remote-vs-proximity'' pairing where, this is the first consideration of this aspect found in the literature. Second, was their studies aim into investigating fingerprint security and usability in a end-to-end encryption context. The study compared Numerical, Visual and Auditory representations. This is a standard setup for similar studies in the literature. QR code verification was also added to the proximity tests.

@@

@@

\subsection*{Methods of comparison performance}
As explained in the definition section ``methods of comparison'' are the ways used to compare fingerprints. Aside from ``Compare-and-Confirm'' (CaC) there is: ``Compare-and-Select'' (CaS) and ``Compare-and-Enter'' (CaE). CaS is the method where one device displays the fingerprint and the other user is provided with a number of options, the user then has to choose the correct value from the list of candidates. If there is no match the user must deny the connection attempt. This was due to concerns that CaC would be ``too easy'' for users leading to complacency and errors\cite{uzun2007usability}. CaE is designed for scenarios where both devices might not have a display, i.e. pairing between a phone and a keyboard. One device displays the checksum and the other device enters the string. The second device then compares the entered string and checks for a match.

@@

One aspect of note from the work by \textbf{Ersin Uzun \textit{et al.}}\cite{uzun2007usability} is their studies context  of  “Secure  device  pairing”  and  thus  some comparison schemes are tested that are not relevant in a fingerprint comparison context, for example “Choose-and-Enter” (A participant chooses a passphrase from an available list and enters it into the other device).  Thus, only relevant comparison schemes and their results have been extracted and compared. \\

\subsection*{Experimentation methodology comparison}
To further look into the validity of previously discussed results it is necessary to asses how the respective studies reached these conclusions. Areas for consideration are scheme entropy, attacker strength and participant factors among others.\\
The starkest limitations of the literature are the range of participant and encoding entropy, where at its worst only 22-bits of entropy were tested. This makes it difficult to directly compare results and needs to be taken into consideration when comparing.
One of the papers this most affects is the early work by \textbf{Hsu-Chun Hsiao, \textit{et al.}} where their highest entropy is 28-bits. This was small even at the time of publication. There is an attempt by the authors' to address this issue in the later stages of the paper where they write \textit{``[...] increasing entropy is not a solution because it sacrifices usability and accuracy. With more entropy, representations will contain longer sequences of characters or more minute details which will lead to increased time and errors during comparisons.''}. Although the assumptions in this statement could be seen as following common sense they fail to consider the implications of of a small entropy on the security of the system.
\\
This leads into the assessment of attacker strength considerations made by each paper. Some papers failed to directly address their attacker strength consideration, but the overall strength has been infered from the changes made to their schemes. For example, if they decided to change a single character in a 40-digit (160-bit) SHA-1 hex digest they are indirectly stating that the attacker can control 39-digits (156-bit). To achieve this the attacker would have to compute $2^{156}$ SHA-1 compressions to find a key match.
In the literature this element ranged from $2^{28}$ to around $2^{242}$, however, this has some relation to the size of the encoding schemes used.
\\
In terms of the studies performed they all considered demographical data when presenting their results. Their average ages were all around $\sim35$ years old with the majority of participants being educated with at least a bachelor degree. This was alongside the equal split between male and female.

One consideration of note is that made by Dechand \textit{et al.} \cite{dechand2016empirical} where they briefly consider medical conditions such as ADHD and reading or visual disorders and they way they affect the comparison's effectiveness. They highlight on a slight reduction in accuracy but due to their small sample size they cannot conclusively validate these results. This is an aspect unique to all other literature, further work would be required to produce conclusive results, thus showing a gap in the literature.

@@

\textbf{TODO: } Do an overall comparison of the sizes of the studies and link to the table produced

@@

>>>>

\subsection*{Encoding scheme performance}
Results 
from the literature consistently show the effectiveness of language based encodings (Words or Sentences). This was shown in \cite{dechand2016empirical}, \cite{tan2017can} and \cite{kainda2009usability} with the accuracy of these schemes all being above 94.00\%. In all cases these were the best scemes in terms of accuracy. The exception to this is the work performed by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study} in 2009 with Words achieving an abnormal accuracy of 63.00\% (Comparisons of methodology will be discussed in later sections).
\\\\
Aside from textual representations were graphical schemes. Examples of schemes assessed were: Random Art\cite{perrig1999hash}, Flag\cite{ellison2003public}, T-Flag\cite{lin2010spate}, Vash\footnote{https://github.com/thevash/vash}, OpenSSH visual host key and Unicorns\footnote{https://unicornify.pictures/} to name a few. Accuracy from these schemes were mixed with ranges as large as 50\% - 94\% in the work by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study}. The only other paper assessing graphical representations was the work of \textbf{Tan \textit{et al.}}\cite{tan2017can} where they also achieved mixed results with accuracies ranging from 46\% to 90\%. With both papers assessing unique schemes, further work is required to validate results.
\\
In terms of usability of graphical schemes, the literature concurred on their high usability. The comparison speed of these schemes were all among the quickest. This is also supported indirectly when comparing to alternative literature that did not consider graphical schemes.\cite{dechand2016empirical}\cite{kainda2009usability}
In terms of research into the performance of graphical schemes, the literature does not contain an extensive review with only two papers including graphical schemes in their comparison. There is also no overlap in the schemes assessed with each paper reviewing a unique set. This is, therefore, a promising candidate for further research.

One unique paper in this research area was the work by \textbf{M. Shirvanian \textit{et al.}}\cite{shirvanian2017pitfalls} produced in the context of secure messaging pairing. This paper was unique for a number of reasons. First was to consideration for ``remote-vs-proximity'' pairing where, this is the first consideration of this aspect found in the literature. There is room for further research with the aim to compare encoding schemes in the context of ``remote-vs-proximity''.
Another unique aspect was the end-to-end encryption context the study was performed under.

@@

Alongside this, it was shown in work by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study} that age and gender do not affect the accuracy of the scheme. However, younger participants were considerably faster. Further more, findings also showed that language comprehension helped in discerning small differences between schemes encoded in Chinese, Japanese, Korean or English. Subsequently, knowledge of the language did not assist in differentiating larger changes in the schemes as these had high accuracy regardless. These were interesting and unique considerations. Further work with these considerations in mind could be tasked with corroborating these findings.

@@

\subsection*{Methods of comparison performance}
As explained in the definition section ``methods of comparison'' are the ways used to compare fingerprints. Aside from ``Compare-and-Confirm'' (CaC) there is: ``Compare-and-Select'' (CaS) and ``Compare-and-Enter'' (CaE). CaS is the method where one device displays the fingerprint and the other user is provided with a number of options, the user then has to choose the correct value from the list of candidates. If there is no match the user must deny the connection attempt. CaS was devised due to concerns that CaC would be ``too easy'' for users leading to complacency and errors\cite{uzun2007usability}. CaE is designed for scenarios where both devices might not have a display, i.e. pairing between a phone and a keyboard. One device displays the checksum and the other device enters the string. The second device then compares the entered string and checks for a match.

@@

One aspect of note from the work by \textbf{Ersin Uzun \textit{et al.}}\cite{uzun2007usability} is their studies context  of  “Secure  device  pairing”  and  thus  some comparison schemes are tested that are not relevant in a fingerprint comparison context, for example “Choose-and-Enter” (A participant chooses a passphrase from an available list and enters it into the other device).  Thus, only relevant comparison schemes and their results were extracted and compared. \\

\subsection*{Experimentation methodology comparison}
To further look into the validity of previously discussed results it is necessary to asses how the respective studies reached these conclusions. Areas for consideration are scheme entropy, attacker strength and participant attributes.

The starkest limitations of the literature are the range of participant and encoding entropy, where at its worst only 22-bits of entropy were tested. This makes it difficult to directly compare results and needs to be taken into consideration when comparing. One of the papers this most affects is the early work by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study}  where their highest entropy is 28-bits. This was small even at the time of publication. There is an attempt by the authors' to address this issue in the later stages of the paper where they write \textit{``[...] increasing entropy is not a solution because it sacrifices usability and accuracy. With more entropy, representations will contain longer sequences of characters or more minute details which will lead to increased time and errors during comparisons.''}. This statement is backed up with no empirical evidence and the authors' fail to consider how the small entropy would affect the overall security of the schemes.

This leads into the assessment of attacker strength considerations made by each paper. Some papers failed to directly address their attacker strength consideration, but the overall strength has been infered from the changes made to their schemes. For example, if they decided to change a single character in a 40-digit (160-bit) SHA-1 hex digest they are indirectly stating that the attacker can control 39-digits (156-bit). To achieve this the attacker would have to compute $2^{156}$ SHA-1 compressions to find a key match.
In the literature this element ranged from $2^{28}$ to around $2^{242}$, however, this has some relation to the size of the encoding schemes used. This is a substantial range that makes it ultimately difficult to confidently compare and confer results. Further work could exclusively look into the affect attacker strength has on the success of attacks.

All of the studies considered demographical data when presenting their results. Their average ages were all around $\sim35$ years old with the majority of participants being educated with at least a bachelor degree. This was alongside the equal split between male and female.

One consideration of note is that made by Dechand \textit{et al.} \cite{dechand2016empirical} where they briefly consider medical conditions such as ADHD and reading or visual disorders and they way they affect the comparison's effectiveness. They highlight on a slight reduction in accuracy but due to their small sample size they cannot conclusively validate these results. This is an aspect unique to all literature. Further work would be required to produce conclusive results. Therfore, highlighting on a gap in the literature.

@@

Hexadecimal and numerical schemes were not included as encoding schemes in the work by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study} (some of the most widespread encoding schemes). This was based on their own claims on similarities to Base32 and ``well-known deficiencies" in the excluded schemes. This point was provided with no further justification or quantified in any way. It is also not consistent with available research, for example, in \textit{``Empirical Study of Textual Key-Fingerprint Representations''}\cite{dechand2016empirical} it was shown numerical representations performed significantly better than that of Base32.

@@

Table \ref{tab:attacker} has been provided to visually compare the differing aspects of the papers' parameters. Clearly it can be seen from the table the large ranges in participant size, entropy and attacker strength.
